HO unification from object language to meta language

### 2024-03-28 paper/main.tex
update proof for progress termination

As a final remark, the \maybeeta detection defined just before is an
over-approximation, in the sense that there exists some terms $t$ considered as
\maybeeta, such that forall substitution $\sigma$, $\sigma t$ is not an
$\eta$-expansion. A small example is $\lambda x. f \appsep A_x \appsep A_x$. On
the other hand, we also point out that the \maybeeta detection spot out
potential $\eta$-expansion for terms that are not in \llambda. For example,
$\lambda x.F \appsep G_x$ is considered as \maybeeta, since we have the
application of term whose argument can reduce to $x$.

@@

We can note that the insertion of these rules for \elpiIn{progress1} do not
prevent the termination of \elpiIn{progress}, since: if a link is activated it
is removed from \linkStore and the recursive call to \elpiIn{progress} will have
a smaller list of links to recurse on. Moreover, link activation only runs
terminating instructions (such as unification). If a link is deduplicated, the
termination of \elpiIn{progress} is still guaranteed since again we reduce
\linkStore and the instructions run by link deduplications are all terminating.
Finally, if a link is neither activated nor deduplicated, i.e. it remains
suspended, then the condition of the \elpiIn{if}-statement of \elpiIn{progress}
succeeds making it to terminate.

@@

$\beta$-reduction problems (\maybebeta) appears any time we deal with a subterm $t
= X t_1 \dots t_n$, where $X$ is flexible and the list $[t_1 \dots t_n]$ in not
in \llambda. This unification problem is not solvable without loss of
generality, since there is not a most general unifier. If we take back the
example given in \cref{sec:nutshell}, the unification $F a = a$ admits two solutions for $F$:
$\rho_1 = \{F \mapsto \lambda x.x\}$ and $\rho_2 = \{F \mapsto \lambda \_.a\}$.
Despite this, it is possible to work with $\maybebeta$ if an oracle provides a
substitution $\rho$ such that $\rho t$ falls again in the \llambda.

>>>>

Finally, the \maybeeta detection defined just before is an
over-approximation, in the sense that there exists some terms $t$ considered as
\maybeeta, such that forall substitution $\sigma$, $\sigma t$ is not an
$\eta$-expansion. A small example is $\lambda x. f \appsep A_x \appsep A_x$.

@@

\begin{lemma}
  Forall list of links \elpiIn{L} and \elpiIn{S}, \elpiIn{progress L _ S _}
  terminates
\end{lemma}

\begin{proof}[Proof sketch]
  The addition of rules for \elpiIn{progres1} complicates the function
  \elpiIn{progress}. We can note, however, that they do not prevent the
  termination of \elpiIn{progress}. 1) If a link is activated it is removed from
  \linkStore and the recursive call to \elpiIn{progress} will have a smaller
  list of links to recurse on. Moreover, link activation only runs terminating
  instructions (such as unification). 2) If a link is deduplicated, the
  termination of \elpiIn{progress} is still guaranteed since again we reduce
  \linkStore and the instructions run by link deduplications are all
  terminating. 3) If a link is neither activated nor deduplicated, i.e. it
  remains suspended, then \linkStore remains unchanged like the substitution;
  therefore, \elpiIn{if (L = L1, S1 = S2)} succeeds and \elpiIn{progress}
  terminates.
\end{proof}

@@

TODO: say that maybe-eta also work in not(llambda)

% On
% the other hand, we also point out that the \maybeeta detection spot out
% potential $\eta$-expansion for terms that are not in \llambda. For example,
% $\lambda x.F \appsep G_x$ is considered as \maybeeta, since we have the
% application of term whose argument can reduce to $x$.

All the previous sections we have dealt with terms in \llambda, however, it is
often possible to work with terms in \maybebeta and wish to unify them. There
are situtation, for example, where the oracle has given  which $\beta$-reduction problems (\maybebeta) appears any
time we deal with a subterm $t = X t_1 \dots t_n$, where $X$ is flexible and the
list $[t_1 \dots t_n]$ in not in \llambda. This unification problem is not
solvable without loss of generality, since there is not a most general unifier.
If we take back the example given in \cref{sec:nutshell}, the unification $F a =
a$ admits two solutions for $F$: $\rho_1 = \{F \mapsto \lambda x.x\}$ and
$\rho_2 = \{F \mapsto \lambda \_.a\}$. Despite this, it is possible to work with
$\maybebeta$ if an oracle provides a substitution $\rho$ such that $\rho t$
falls again in the \llambda.
